A numerical study of the solution of X-Mode equations around the hybrid resonance

### 2015-04-29 report_clean/responce_reviewers.tex

-p 6: $Ex$ changed

-p 8: our code, changed

 - p.8: fixed section number

 - p.8 eq (12): added $\nu=0$ + explanation

 - p.8: added the Airy equation. Additionally added a remark on the behaviour of the Airy function to justify the choice of the right boundary 
 condition $\partial_x E=0$.

 -  p9, Figure 3: fixed the captions. This had been done for other figures as well. The values $E^{c}_{x,y}$ had been defined in the caption.

 -p9 on top
  The title of Section 4.1.2 had been changed to Solution of a Frequency-Domain Problem with Resonance

 - p9 Equation (15,16): changed

 - p9 after (17): we would like to assume that it is $H^2$-norm that is bounded, for using the inequality (19). 
  \mrev{TODO: While this assumption is fairly strong, it is not unsubstantial: for sufficiently smooth $\alpha,\delta$,  }

 - p10 before and after Table 1 : fixed

 -- p11 Figure 5:
Agree, fixed. We added a remark in a caption to Figure 4 as well.

- p12 Section 4.3: this explanation had been moved to the introduction

- p12 Section 4.3.1

  It is true, but in this case there is no resonance, since $\delta(x)\equiv 0$. Hence $E_x=-\frac{i\delta}{\alpha}E_y\equiv 0$, 
  and $E_y$ solves the Airy equation in the frequency domain. We added a small to Section 4.3.1 to underline this.

- p12 Equation (23)
We changed it to $e$

- p14 Figures 8 and 9
Added the captions  
- p15 in Conclusion
  Indeed $P1$!

 Reviewer 2:

 • p2: In this case Ex is not square integrable: confirm that it is when ν is not zero and that
the singularity appears for ν = 0.

We tried to change the wording in the introduction to make it more clear what is a hybrid resonance. 
The singularity appears for $\nu=0$, when the diagonal part of the dielectric tensor vanishes and non-diagonal part does not.

• p6, 3.1. It is not clear how the Yee scheme, which is a finite difference scheme, is linked to the
finite element scheme developed in the previous section. This should be made more explicit.

We employed the Yee scheme to discretize the problem in the time domain, and the finite element scheme solely for the frequency domain, 
and considered them independently. We added a remark on that in the very beginning of Section 3.

• p10: mesh multiple times finer. It should be said exactly how many times.

We added a remark that it was at least two times for very fine meshes, and more for coarser ones.

### 2015-04-29 report_clean/numerics.tex
final changes. To-Do: more changes

\subsection{Frequency Domain Problem}
\label{sec:freq_dep}
\subsubsection{Validity of the Implementation}
To check the validity of the code, we first perform a numerical experiment with (formally chosen) parameters:
\begin{align}
\label{eq:parameters}
\alpha(x)=x^2+1,\qquad \delta(x)=\left(\alpha^2+x\alpha\right)^{\frac{1}{2}}, \qquad \mrev{\nu=0}.
\end{align}
\mrev{It can be shown that $E_{y}$ that solves \mrev{(\ref{eq:main_frequency_domain_general})} in 1D, with the Fourier variable $\theta=0$ (see (\ref{vf1dcase}), 
and $\nu=0$, satisfies the Airy equation, see also Remark \ref{remark:other},
\begin{align*}
 E_y''-xE_y=0,\qquad x\in [-L,\; H].
\end{align*}
One of the solutions of this equation, an Airy function $\Ai(x)$, see \cite[Chapter 10.4]{abramowitz_stegun}, decays rapidly as $x\rightarrow \infty$. 
Hence, choosing $H$ sufficiently large \mrev{TODOOO(the actual value used in the experiments $H=10$)} and the boundary conditions as 
\begin{align}
\label{eq:bcs_airy}
\partial_{1}E_{y}(-L)+2iE_{y}(-L)=2i\Ai(-L)+\Ai'(-L),\qquad 
\partial_{1}E_{y}(H)=0\approx Ai(H),
\end{align}
where $\Ai(x)$ is an Airy function, we ensure that $E_{y}$ that solves the above problem is well approximated by $\Ai(x)$, and $E_{x}$ by $-i\frac{\delta(x)}{\alpha(x)}\Ai(x)$. 
The well-posedness of the respective variational formulation for $\nu=0$ is due to Remark \ref{remark:other}. }

@@

          \addplot[mark=*,cyan, mark size=1pt] table {pics_frequency_domain/data_airy/E2L2.dat};
        \addlegendentry{$\|E_{y}-E_{y}^{c}\|_{L_{2}}$};  
        \addplot[dotted] table{pics_frequency_domain/data_airy/E_h2.dat};
         \addlegendentry{$O(h^2)$};
         \addplot[dashed] table{pics_frequency_domain/data_airy/H.dat};
         \addlegendentry{$O(h)$};
    \end{loglogaxis}
    \end{tikzpicture}
    \caption{Convergence rates for the problem with parameters (\ref{eq:parameters}) with the boundary condition (\ref{eq:bcs}).}
    \label{fig:conv_rate}
\end{figure}
\subsubsection{Solution of $X$-Mode Problem}
%\paragraph{Convergence}
Let us consider the case of the resonance, more precisely, we choose sufficiently smooth
$\alpha,\delta$, s.t. $\alpha(0)=0$ and $\delta(0)\neq 0$, and the solvability conditions 
of Lemma \ref{lemma:well_posedness}  are satisfied. 
For simplicity, let us assume 
\begin{align}
\label{eq:cond}
 \alpha(x)=-x \text{  in some neighbourhood of $0$ }.
\end{align}
Given $\mathbf{V}_{h}=P_{h}^{1}\times P_{h}^{1}$, with $P_{h}^{1}$ consisting of piecewise-linear (hat) functions, 
we look for a ratio $h(\nu)$ that would ensure the bound on the absolute error 
\begin{align}
\label{eq:problem1}
\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}<\epsilon,
\end{align}
given a fixed value of $\epsilon>0$ and $\nu\rightarrow 0$. W.l.o.g. here we assume that $\nu>0$.

We use the following ingredients:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq C\nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V},\; C>0.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness} and is valid for $\nu\rightarrow 0$.   
\item The form of the exact solution to the problem (\ref{vf1dcase}), see \cite{Despres_2014},
\begin{align}
\label{eq:exact}
 E_x^{\nu}=-iE_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f^{\nu}(x)}{\alpha(x)+i\nu}, \; \text{for some }f^{\nu}(x)\in L_{2}(\Omega).
\end{align}
We additionally assume that $E_y^{\nu}(x)$ is sufficiently smooth and $\|(E_y^{\nu})''\|_{H_{2}(\Omega)}$
can be bounded independently of $\nu$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:interp}
\begin{split}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2\left|v''\right|_{L_{2}(\Omega)},\; C>0,\\
 \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
 \end{split}
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$. 
\end{itemize}
The estimate (\ref{eq:cea}) combined with (\ref{eq:interp}) results in, for some $C>0$,
\begin{align}
  \label{eq:final_est}
  \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}%\leq C_1\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L_{2}}+C_2\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}}
  \leq C\nu^{-1}h+C\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}},
\end{align}
where we used an assumption that the $L_2$-norm of $\left(E_y^{\nu}(x)\right)''$ can be bounded independently of $\nu$.

For $E_x^{\nu},\; f^{\nu}(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{(f^{\nu})''}{\alpha+i\nu}-2\frac{(f^{\nu})'\alpha'}{(\alpha+i\nu)^2}+\frac{f^{\nu}\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\bealn
 \left|\frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
 % &=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. After inserting this into (\ref{eq:final_est}) we obtain
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2+C\nu^{-1}h.
\een
From this it thus follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. 
This prediction is very severe and is due to the singular nature of the problem under consideration.

Let us check numerically whether this holds true. 
To do so, we conduct the  numerical experiment with parameters for the problem as given in Table \ref{tab:parameters}.
\begin{table}[htb!]
\begin{tabular}{l|l}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{lr}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{lr}
0, & x\leq -10,\\
4/30x+4/3,& -10<x\leq 5,\\
2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance.}
\label{tab:parameters}
\end{table}
Denoting by $E_{x}^{h}$ the solution $E_{x}(x)$ computed on the mesh with a width $h$, and by $E_{x}^{c}$ the solution computed on a mesh multiple times finer, let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{h'}-E_{x}^{c}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/h_nu.pdf}
&
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/res_sol.pdf}
\end{tabular}
\caption{In the left figure the dependence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated. 
In the right figure we show the computed solutions for the problem with parameters in Table \ref{tab:parameters} for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that the estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for given values of $\epsilon$ and for a chosen range of $\nu>0$.
\begin{remark}
It can be shown that $\|E^{\nu}_{x}\|_{L_{2}(\Omega)}\leq \frac{C}{\sqrt{\nu}},\; C>0$, 
and thus the relative error control
\bealn
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}}{\|E^{\nu}_{x}\|_{L_{2}(\Omega)}}\leq \epsilon
\eealn
 is ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{4}}$.
\end{remark}

In Figure \ref{fig:small_nu} we demonstrate the dependence of the condition number of the matrix of the system (\ref{eq:simple_system}) 
on $\nu$, for several values of $h$.
Remarkably, for $\nu=0$ the computed matrices are not singular. We do not know the exact reason for this.
As an additional illustration, we compare the solutions for very small $\nu$ and $\nu=0$ for different meshes in Fig.~\ref{fig:small_nu}. 
\begin{figure}[htb!]
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/fig_cond_num.pdf}&
 \includegraphics[height=0.32\textwidth]{pics_frequency_domain/ey_resonance.pdf}
 \end{tabular}
 \caption{
 The left plot demonstrates the dependence of the condition number of the system (\ref{eq:simple_system}) on $\nu$, for different values of $h$ (the condition 
 number for $\nu=0$ is not shown, however, for all values of $h$ as shown in the plot, the matrix was non-singular even for $\nu=0$). 
 The right plot shows $E_{y}^{\nu}$ computed on different meshes for small $\nu$. It can be seen that for $h=10^{-2}$ and $h=10^{-3}$ the solutions 
 computed for $\nu=0$ are almost indistinguishable.}
 %It can be seen that for meshes that are coarse 
 %with respect to $\nu$ (e.g. $\nu=10^{-4}, \; h=10^{-2}$, or $\nu=0,\; h=10^{-2}$), the computed solution seems to be only piecewise-continuous in $\nu$, 
 %while for finer meshes and $\nu>0$ the solution seems to be smooth. }
 \label{fig:small_nu}
\end{figure}
%\begin{table}[ht!]
%\begin{tabular}{c|ccccccccc}
%\diagbox[width=10em]{$h$}{$\nu$}& 100 & 10 & 1 & 0.1 & 1e-2 & 1e-3 & 1e-4 & 1e-8 & 0\\
% \hline
%1 & 3.9634&4.3787 &4.987&36.949&42.394& 42.757 &42.79&42.793&42.793\\
%\hline
%0.5&      3.9931& 6.87&48.473 & 168.41  &207.34  & 209.63 & 209.83 & 209.85&  209.85\\
%\hline
% 0.1&      16.09  &     159.56  &     1146.1   &    8304.1   &     19169  &      20270   &     20344   &     20352   &     20352\\
% \hline
%0.05&      64.046  &     637.18  &     4686.2   &     38876 &  1.4e5 &  1.58e5 &   1.58e5 &  1.58e5 &  1.58e5\\
%\hline
% 0.01&       1600    &   15921 & 1.2e5  & 1.1e6 & 8.3e6  & 1.8e7   & 1.93e7  & 1.93e7  & 1.93e7\\
%\end{tabular}
%\caption{The condition number of the matrix of the system (\ref{eq:simple_system}) for different values of $\nu$ and $h$. }
%\label{tab:cond_number}
%\end{table}

>>>>

\subsection{Frequency Domain Problem}
\label{sec:freq_dep}
\subsubsection{Validity of the Implementation}
To check the validity of the code, we first perform a numerical experiment with (formally chosen) parameters:
\begin{align}
\label{eq:parameters}
\alpha(x)=x^2+1,\qquad \delta(x)=\left(\alpha^2+x\alpha\right)^{\frac{1}{2}}, \qquad \mrev{\nu=0, \qquad \theta=0.}
\end{align}
\mrev{It can be shown that $E_{y}$ that solves (\ref{eq:main_frequency_domain_general}) in 1D, with the Fourier variable $\theta=0$ (see (\ref{vf1dcase}), 
and $\nu=0$, satisfies the Airy equation, see also Remark \ref{remark:other},
\begin{align*}
 E_y''-xE_y=0,\qquad x\in [-L,\; H].
\end{align*}
One of the solutions of this equation, an Airy function $\Ai(x)$, see \cite[Chapter 10.4]{abramowitz_stegun}, is analytic and decays rapidly as $x\rightarrow \infty$. 
Hence, choosing $H$ sufficiently large \mrev{TODOOO(the actual value used in the experiments $H=10$, where $\Ai'(10)\approx -3.5e-10$)} and the boundary conditions as 
\begin{align}
\label{eq:bcs_airy}
\partial_{1}E_{y}(-L)+2iE_{y}(-L)=2i\Ai(-L)+\Ai'(-L),\qquad 
\partial_{1}E_{y}(H)=0\approx \Ai'(H),
\end{align}
where $\Ai(x)$ is an Airy function, we ensure that $E_{y}$ that solves the above problem is well approximated by $\Ai(x)$, and $E_{x}$ by $-i\frac{\delta(x)}{\alpha(x)}\Ai(x)$. 
The well-posedness of the respective variational formulation for $\nu=0$ is due to Remark \ref{remark:other}. }

@@

          \addplot[mark=*,cyan, mark size=1pt] table {pics_frequency_domain/data_airy/E2L2.dat};
        \addlegendentry{$\|E_{y}-E_{y}^{c}\|_{L_{2}}$};  
        \addplot[dotted] table{pics_frequency_domain/data_airy/E_h2.dat};
         \addlegendentry{$O(h^2)$};
         \addplot[dashed] table{pics_frequency_domain/data_airy/H.dat};
         \addlegendentry{$O(h)$};
    \end{loglogaxis}
    \end{tikzpicture}
    \caption{Convergence rates for \mrev{the problem (\ref{eq:main_frequency_domain}) discretized as (\ref{eq:simple_system})} 
    with the parameters (\ref{eq:parameters}) and the \mrev{boundary condition (\ref{eq:bcs_airy})}. \mrev{Here $E_{y}^{c}(x)=\Ai(x)$ and 
    $E_{x}^{c}(x)=-i\frac{\delta(x)}{\alpha(x)}\Ai(x)$.} }
    \label{fig:conv_rate}
\end{figure}
\subsubsection{Solution of a Frequency-Domain Problem with Resonance}
%\paragraph{Convergence}
Let us consider the case of the resonance, more precisely, we choose sufficiently smooth
$\alpha,\delta$, s.t. $\alpha(0)=0$ and $\delta(0)\neq 0$, and the solvability conditions 
of Lemma \ref{lemma:well_posedness}  are satisfied. 
For simplicity, let us assume 
\begin{align}
\label{eq:cond}
 \alpha(x)=-x \text{  in some neighbourhood of $0$ }.
\end{align}
Given $\mathbf{V}_{h}=P_{h}^{1}\times P_{h}^{1}$, with $P_{h}^{1}$ consisting of piecewise-linear (hat) functions, 
\mrev{and $E^{\nu}_{x},\;E^{\nu,h}_{x}$ being correspondingly the exact solution and the solution computed with the help $\mathbf{V}_h$, }
we look for a ratio $h(\nu)$ that would ensure the bound on the absolute error 
\begin{align}
\label{eq:problem1}
\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}<\epsilon,
\end{align}
given a fixed value of $\epsilon>0$ and $\nu\rightarrow 0$. W.l.o.g. here we assume that $\nu>0$.

We use the following ingredients:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here \mrev{$\mathbf{E}^{\nu}$ is the exact solution, 
 $\mathbf{E}^{\nu,h}$ is the computed solution}, and $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{\nu,h}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}^{\nu}-\mathbf{v}\|_{V}
 \leq C\nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}^{\nu}-\mathbf{v}\|_{V},\; C>0.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness} and is valid for $\nu\rightarrow 0$.   
\item The form of the exact solution to the problem (\ref{vf1dcase}), see \cite{Despres_2014},
\begin{align}
\label{eq:exact}
 E_x^{\nu}=-iE_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f^{\nu}(x)}{\alpha(x)+i\nu}, \; \text{for some }f^{\nu}(x)\in L_{2}(\Omega).
\end{align}
We additionally assume that $E_y^{\nu}(x)$ is sufficiently smooth and $\|(E_y^{\nu})''\|_{H_{2}(\Omega)}$
can be bounded independently of $\nu$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:interp}
\begin{split}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2\left|v''\right|_{L_{2}(\Omega)},\; C>0,\\
 \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
 \end{split}
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$. 
\end{itemize}
The estimate (\ref{eq:cea}) combined with (\ref{eq:interp}) results in, for some $C>0$,
\begin{align}
  \label{eq:final_est}
  \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}%\leq C_1\nu^{-1}h\|\left(E_y^{\nu}\right)''\|_{L_{2}}+C_2\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}}
  \leq C\nu^{-1}h+C\nu^{-1}h^2\|\left(E_x^{\nu}\right)''\|_{L_{2}},
\end{align}
where we used an assumption that the $L_2$-norm of $\left(E_y^{\nu}(x)\right)''$ can be bounded independently of $\nu$.

For $E_x^{c,\nu},\; f^{\nu}(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{(f^{\nu})''}{\alpha+i\nu}-2\frac{(f^{\nu})'\alpha'}{(\alpha+i\nu)^2}+\frac{f^{\nu}\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\bealn
 \left|\frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
 % &=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. After inserting this into (\ref{eq:final_est}) we obtain
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2+C\nu^{-1}h.
\een
From this it thus follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. 
This prediction is very severe and is due to the singular nature of the problem under consideration.

Let us check numerically whether this holds true. 
To do so, we conduct the  numerical experiment with parameters for the problem (\ref{eq:main_frequency_domain} as given in Table \ref{tab:parameters}.
\begin{table}[htb!]
\begin{tabular}{l|l}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{lr}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{lr}
0, & x\leq -10,\\
4/30x+4/3,& -10<x\leq 5,\\
2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance.}
\label{tab:parameters}
\end{table}
\mrev{Recall that we choose $\theta=0$. }
Denoting by \mrev{$E_{x}^{\nu,h}$ the solution $E_{x}(x)$} computed on the mesh with a \mrev{ width $h$ with 
the absorption $\nu$, and by $E_{x}^{\nu}$} the solution computed on a mesh multiple times finer (\mrev{more precisely, the reference mesh
was typically 10-20 times finer, with the exception when the computation was done for very small values of $h$ where twice finer meshes for comparison were used}), let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{\nu,h'}-E_{x}^{\nu}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/h_nu.pdf}
&
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/res_sol.pdf}
\end{tabular}
\caption{We consider the problem (\ref{eq:main_frequency_domain}) with the parameters 
defined in Table \ref{tab:parameters}. In the left figure the dependence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated.  
In the right figure we show the computed solutions for this problem (\mrev{for brevity denoted as $E_{x,y}^{\nu}$}) for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that the estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for given values of $\epsilon$ and for a chosen range of $\nu>0$.
\begin{remark}
It can be shown that $\|E^{\nu}_{x}\|_{L_{2}(\Omega)}\leq \frac{C}{\sqrt{\nu}},\; C>0$, 
and thus the relative error control
\bealn
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}}{\|E^{\nu}_{x}\|_{L_{2}(\Omega)}}\leq \epsilon
\eealn
 is ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{4}}$, \mrev{$\beta_{\epsilon}>0$.}
\end{remark}

In Figure \ref{fig:small_nu} we demonstrate the dependence of the condition number of the matrix of the system (\ref{eq:simple_system}) 
on $\nu$, for several values of $h$.
Remarkably, for $\nu=0$ the computed matrices are not singular. We do not know the exact reason for this.
As an additional illustration, we compare the solutions for very small $\nu$ and $\nu=0$ for different meshes in Fig.~\ref{fig:small_nu}. 
\begin{figure}[htb!]
\begin{tabular}{cc}
\includegraphics[height=0.32\textwidth]{pics_frequency_domain/fig_cond_num.pdf}&
 \includegraphics[height=0.32\textwidth]{pics_frequency_domain/ey_resonance.pdf}
 \end{tabular}
 \caption{
 We consider the problem (\ref{eq:main_frequency_domain}) with the parameters 
defined in Table \ref{tab:parameters}. The left plot demonstrates the dependence of the condition number of the system (\ref{eq:simple_system}) on $\nu$, for different values of $h$ (the condition 
 number for $\nu=0$ is not shown, however, for all values of $h$ as shown in the plot, the matrix was non-singular even for $\nu=0$). 
 The right plot shows \mrev{$E_{y}^{\nu,h}$} computed on different meshes for small $\nu$. It can be seen that for $h=10^{-2}$ and $h=10^{-3}$ the solutions 
 computed for $\nu=0$ are almost indistinguishable.}
 %It can be seen that for meshes that are coarse 
 %with respect to $\nu$ (e.g. $\nu=10^{-4}, \; h=10^{-2}$, or $\nu=0,\; h=10^{-2}$), the computed solution seems to be only piecewise-continuous in $\nu$, 
 %while for finer meshes and $\nu>0$ the solution seems to be smooth. }
 \label{fig:small_nu}
\end{figure}
%\begin{table}[ht!]
%\begin{tabular}{c|ccccccccc}
%\diagbox[width=10em]{$h$}{$\nu$}& 100 & 10 & 1 & 0.1 & 1e-2 & 1e-3 & 1e-4 & 1e-8 & 0\\
% \hline
%1 & 3.9634&4.3787 &4.987&36.949&42.394& 42.757 &42.79&42.793&42.793\\
%\hline
%0.5&      3.9931& 6.87&48.473 & 168.41  &207.34  & 209.63 & 209.83 & 209.85&  209.85\\
%\hline
% 0.1&      16.09  &     159.56  &     1146.1   &    8304.1   &     19169  &      20270   &     20344   &     20352   &     20352\\
% \hline
%0.05&      64.046  &     637.18  &     4686.2   &     38876 &  1.4e5 &  1.58e5 &   1.58e5 &  1.58e5 &  1.58e5\\
%\hline
% 0.01&       1600    &   15921 & 1.2e5  & 1.1e6 & 8.3e6  & 1.8e7   & 1.93e7  & 1.93e7  & 1.93e7\\
%\end{tabular}
%\caption{The condition number of the matrix of the system (\ref{eq:simple_system}) for different values of $\nu$ and $h$. }
%\label{tab:cond_number}
%\end{table}
